\begin{abstract}
\noindent This paper examines whether predicted crop yields causally affect conflict at the subnational level across six Sub-Saharan African countries---Ethiopia, Malawi, Mali, Nigeria, Tanzania, and Uganda---over the period 2010--2024. A persistent identification problem in the climate--conflict literature is that observed yields are endogenous: conflict itself destroys agricultural production. To address this, I generate satellite-based machine-learning predictions of maize yield for 926 Admin-2 districts using XGBoost models trained on Living Standards Measurement Study (LSMS-ISA) survey data and a rich set of remote-sensing features, including vegetation indices, temperature, and growing-degree days. These predicted yields, constructed entirely from exogenous environmental inputs, serve as the primary regressor in a two-way fixed-effects (TWFE) panel framework controlling for Admin-2 and year fixed effects and time-varying district population. Across the preferred linear specifications, I find no statistically significant relationship between predicted yield and the 3-month forward conflict count. However, a log-log TWFE OLS specification yields a positive and significant elasticity ($\hat{\beta} = 0.175$, $p = 0.022$), suggesting that a 1\% increase in predicted yield is associated with a 0.175\% increase in log conflict events---a pattern more consistent with a predation or resource-contestation mechanism than with an opportunity-cost channel. Robustness checks across 6-month and 12-month conflict windows, binary conflict outcomes, and fatality counts do not alter this conclusion. Instrumental variables (2SLS) estimates, which instrument observed yields with predicted yields, are uninformative due to weak first stages ($F < 2$ in all specifications), reflecting the sparse spatial overlap between LSMS survey years and the full panel. Results should be interpreted cautiously given near-zero yield model predictive accuracy for Nigeria and Uganda.
\end{abstract}
