% ══════════════════════════════════════════════════════════════════════════════
\section{Data}
\label{sec:data}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Panel Structure and Unit of Analysis}

The analysis is based on an Admin-2 district $\times$ year panel covering seven Sub-Saharan African countries---Burkina Faso, Ethiopia, Mali, Malawi, Niger, Nigeria, and Tanzania---over the period 2010--2024. The unit of analysis is the second-level administrative division (Admin-2), corresponding to departments, woredas, cercles, traditional authorities, local government areas, or districts depending on the country. Boundaries are taken from the Global Administrative Areas (GADM) Level-2 dataset, which provides consistent subnational polygon definitions aligned with the yield prediction inputs described in Section~\ref{sec:empirical}.

The panel comprises 18,096 district-year observations across 1,338 distinct Admin-2 units. Table~\ref{tab:panel_summary} reports panel dimensions and key variable summaries by country. Nigeria contributes the largest share (9,101 observations; 715 local government areas), while Niger contributes the smallest (501 observations; 35 departments). Uganda is excluded from all analyses: the yield prediction model was trained on GROW-Africa administrative records from seven countries that do not include Uganda, and leave-one-country-out (LOCO) validation confirms that the model has no out-of-distribution predictive power for Ugandan districts ($R^2 < 0$).

\begin{table}[H]
\centering
\caption{Panel Summary by Country}
\label{tab:panel_summary}
\begin{threeparttable}
\begin{tabular}{lrrrrrr}
\toprule
Country & ISO3 & Districts & Obs. & Mean Conflict\textsuperscript{a} & \% Zero Conflict & Mean Yield (kg/ha) \\
\midrule
Burkina Faso &BFA &  45 &   675 & 16.0 & 64.3 & 1,353 \\
Ethiopia     &ETH &  79 & 1,185 & 22.0 & 71.2 & 2,430 \\
Mali         &MLI &  50 &   750 & 30.2 & 53.6 & 1,492 \\
Malawi       &MWI & 247 & 3,705 &  0.4 & 96.5 & 1,876 \\
Niger        &NER &  35 &   501 &  6.2 & 72.3 & 1,180 \\
Nigeria      &NGA & 715 & 9,101 &  1.6 & 79.3 & 1,702 \\
Tanzania     &TZA & 167 & 2,179 &  0.2 & 95.4 & 1,589 \\
\midrule
\textbf{Total} & & \textbf{1,338} & \textbf{18,096} & \textbf{3.6} & \textbf{82.4} & \textbf{1,674} \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item[a] Mean ACLED conflict events per district-year in the 3-month post-harvest forward window.
\item Mean yield is the panel mean of ML-predicted maize yield (kg/ha).
\item Malawi's 247 Admin-2 units correspond to Traditional Authorities (sub-district); population is assigned at the parent-district level to match the administrative granularity of the population source.
\end{tablenotes}
\end{threeparttable}
\end{table}

\subsection{Conflict Data}

Conflict outcome data are drawn from the Armed Conflict Location and Event Data Project (ACLED), a geocoded database of political violence and protest events that records individual incidents with latitude, longitude, date, event type, actor information, and fatality counts \citep{Raleigh_etal2010}. The full Africa extract covering 2010--2024 is used; raw events are spatially joined to GADM Level-2 polygons and aggregated to district-year counts.

The primary outcome variables are conflict event counts in forward windows following the estimated district harvest month:
\begin{itemize}
  \item \textbf{3-month window} (preferred): total ACLED events in the three calendar months beginning with the harvest month.
  \item \textbf{6-month and 12-month windows}: analogous measures over longer post-harvest horizons, used as robustness checks.
  \item \textbf{Fatalities (3-month)}: total recorded fatalities in the 3-month window, used in an alternative Poisson specification.
\end{itemize}

The forward window anchors the outcome to post-harvest timing so as to capture conflict responses to the yield realisation rather than the reverse. Harvest months are assigned at the Admin-2 level using the GROW-Africa administrative yield database: for countries where sub-national harvest timing varies (Ethiopia, Burkina Faso), the harvest month is the area-weighted mode across administrative units within each zone; for countries with nationally uniform harvest timing, the national modal harvest month is assigned to all districts (ETH: January; BFA, MLI, NER: December; MWI: May; NGA: October; TZA: July).

Conflict event counts are constructed from a complete district $\times$ month grid spanning January 2010 to December 2024. Months recording zero events are explicitly represented, which ensures that forward-window sums correctly reflect the full calendar rather than only months with positive event counts. This is a non-trivial data engineering choice: summing over sparse event-only records would artefactually truncate long-horizon windows when conflict is rare.

A salient distributional feature of the conflict data is the high share of zero district-years: 82.4\% of observations record zero ACLED events in the 3-month window. Zero-inflation is particularly pronounced in Tanzania (95.4\%) and Malawi (96.5\%), while Mali records the lowest zero share (53.6\%). This distributional property motivates the use of Poisson QMLE alongside OLS and the log$(1+x)$ transformation discussed in Section~\ref{sec:empirical}.

\subsubsection{Conflict Type Breakdown}

ACLED classifies events into five broad categories. The analysis uses four: battles, riots, violence against civilians, and explosions/remote violence. These categories have distinct behavioural interpretations. Battles are confrontations between armed groups and are more likely to reflect strategic territorial competition. Riots are collective protest events in which civilian mobilisation predominates. Violence against civilians encompasses deliberate targeting of non-combatants by armed actors. Explosions and remote violence (including IED attacks, airstrikes, and artillery) reflect technologically sophisticated armed activity. The distinction between deliberate targeting events (violence against civilians, explosions) and collective mobilisation events (riots) is directly relevant to separating the resource-predation and grievance channels of the yield--conflict relationship, as discussed in Sections~\ref{sec:theory} and~\ref{sec:results}.

\subsection{Yield Prediction Model and Satellite Data}

Ground-truth yield observations for model training are drawn from the GROW-Africa administrative yield database, a compilation of official ministry-reported maize yield records at the Admin-2 level. The training dataset contains 3,363 observations across 204 administrative units spanning 2000--2023, covering all seven countries in the analysis.

Yield predictions are generated using a global XGBoost model \citep{ChenGuestrin2016} estimated on 210 features derived from the following satellite and environmental sources:

\begin{itemize}
  \item \textbf{Optical vegetation indices}: Enhanced Vegetation Index (EVI), Normalised Difference Vegetation Index (NDVI), Green Chlorophyll Vegetation Index (GCVI), and Leaf Area Index (LAI) from Landsat and Sentinel-2 composites. These indices capture canopy biomass and photosynthetic vigour across diverse cropping systems \citep{BurkeLobell2017}.
  \item \textbf{Synthetic Aperture Radar (SAR)}: Sentinel-1 VV and VH backscatter at 10\,m resolution, providing all-weather structural information complementary to optical indices during cloudy growing seasons.
  \item \textbf{Land Surface Water Index (LSWI)}: a shortwave-infrared index sensitive to vegetation water content and soil moisture status.
  \item \textbf{Climate}: Monthly precipitation from CHIRPS and temperature from ERA5, used to construct growing degree days (GDD) and killing degree days (KDD) \citep{Schlenker_Roberts2009}.
  \item \textbf{Soil properties}: iSDA soil texture, organic carbon, and pH at multiple depths.
  \item \textbf{Terrain}: SRTM elevation, slope, and aspect.
\end{itemize}

All features are aggregated to 12-month seasonal composites relative to each district's harvest month. The model is evaluated using spatial leave-one-admin1-out cross-validation. The global XGBoost achieves a spatial $R^2$ of 0.35 and a stacked ensemble $R^2$ of 0.39---a substantial improvement over earlier single-sensor approaches \citep{LobellEtal2015}. Predictions are generated for all 1,338 GADM Level-2 districts over 2010--2024. The key regressor in all regression specifications is $\log(\hat{y}_{it})$, the natural log of the district-year predicted maize yield in kg/ha.

\subsection{Agro-Ecological Classification}

A secondary analytical dimension is the distinction between \textit{Sahel} and \textit{Non-Sahel} zones, defined at the country level:

\begin{itemize}
  \item \textbf{Sahel zone}: Burkina Faso, Mali, Niger. These countries lie predominantly within the semi-arid Sahel belt (annual rainfall 200--600\,mm), characterised by low and highly variable rainfall, rain-fed cereal systems, and agro-pastoral livelihoods. Armed group activity in this region frequently involves coercive extraction of agricultural surplus through market taxation, cattle raiding, and territorial control of productive zones \citep{Berman_etal2017}.
  \item \textbf{Non-Sahel zone}: Ethiopia, Malawi, Nigeria, Tanzania. These countries span highland, humid tropical, and Guinea Savanna environments with higher mean rainfall and greater yield potential.
\end{itemize}

This classification is a country-level proxy and does not demarcate fine-grained agro-ecological boundaries within countries. Its analytical purpose is to test whether the direction of the yield--conflict elasticity differs systematically between production environments characterised by contrasting conflict mechanisms, as motivated in Section~\ref{sec:theory}.

\subsection{Population Data}

District-level population counts by year are derived from WorldPop 100\,m gridded rasters summed to GADM Level-2 boundaries. The data span 2010--2020; values for 2021--2024 are carried forward from 2020. Population enters regressions as $\log(\text{pop}_{it})$ and controls for the size of the at-risk population in each district. Population is successfully matched to 98.7\% of district-year observations; the remaining 1.3\% (concentrated in a small number of administrative units in Burkina Faso, Niger, and Tanzania with no counterpart in the population source) are excluded from the regression sample.

\subsection{Summary Statistics}

Table~\ref{tab:sumstats} reports summary statistics for the full panel. The conflict distribution is highly right-skewed: the median district-year records zero events, the 90th percentile is 3, and the maximum reaches 1,834. Predicted yield ranges from 386 to 4,463\,kg/ha, with the log specification (mean 7.27, SD 0.42) exhibiting a well-behaved distribution suitable for the log-log regression.

\begin{table}[H]
\centering
\caption{Summary Statistics --- Full Panel}
\label{tab:sumstats}
\begin{threeparttable}
\begin{tabular}{lrrrrrrr}
\toprule
Variable & $N$ & Mean & SD & p10 & Median & p90 & Max \\
\midrule
Conflict events (3-month)             & 18,096 &  3.6  &  36.5 & 0    &  0    &  3    & 1,834 \\
Conflict events (6-month)             & 18,096 &  7.0  &  71.6 & 0    &  0    &  6    & 3,599 \\
Conflict events (12-month)            & 18,096 & 13.3  & 131.1 & 0    &  0    & 11    & 6,614 \\
Fatalities (3-month)                  & 18,096 & 14.0  & 222.7 & 0    &  0    &  2    &11,834 \\
$\log(1+\text{conflict events, 3-mo.})$ & 18,096 &  0.35 &  0.70 & 0    &  0    & 1.39  &  7.52 \\
Predicted yield (kg/ha)               & 18,096 & 1,674 &   643 & 844  & 1,634 & 2,533 & 4,463 \\
Log predicted yield                   & 18,096 &  7.27 &  0.42 & 6.74 &  7.40 &  7.83 &  8.40 \\
Log population                        & 17,889 & 10.30 &  1.31 & 8.72 & 10.30 & 11.89 & 13.38 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\footnotesize
\item $N = 17{,}889$ for log population following exclusion of 207 district-years with unmatched population data.
\item Conflict windows are forward-looking from the district-level estimated harvest month.
\item Log predicted yield is $\log(\hat{y}_{it})$ where $\hat{y}_{it}$ is in kg/ha.
\end{tablenotes}
\end{threeparttable}
\end{table}
